Collective excitations in competing phases in two and three dimensions

### 2024-01-22 paper/main.tex
paper text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                  Abstract                                                 %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
    We investigate the superconducting (SC), charge-density wave (CDW), and antiferromagnetic (AFM) phases 
    in the extended Hubbard model at zero temperature and half-filling.
    We employ the iterated equations of motion approach \cite{Kalthoff17,bleicker18} to compute the two-particle Green's functions and by extension, the corresponding spectral densities.
    This renders comprehensive analysis of the behavior of collective excitations possible as the model's parameters are changed across phase transitions.
    We identify the well-known amplitude (Higgs) and phase (Leggett) modes within the superconducting phase and observe a similar excitation in the CDW phase which shifts towards zero energy as the system approaches the phase transition to the SC phase.
    In the CDW phase, close to the phase transition to the AFM phase, we find a collective mode that does not change significantly and another mode that becomes soft as the phase boundary is approached.
\end{abstract}

@@

In this article, we will first employ a mean-field approximation to the interaction terms 
and then use the methodology from the so-called iterated equations of motion approach (iEoM),
which has already seen success in the handling of interaction quenches \cite{uhrig09,hamerla13,hamerla14,bleicker18}.
Here, one commutes operators from a suitable operator basis with the Hamiltonian and adds the newly occurring terms to it.
Naturally, a truncation is necessary for most practical applications, however, the approximation becomes better the more terms are included within the basis.
The functionality of this method has been compared to the density matrix formalism and has proved to be considerably more accurate \cite{Kalthoff17}.
Additionally, we will show an explicit way to obtain various Green's functions from the aforementioned methodology.
By extension, we also obtain the spectral functions of the investigated systems and discuss the excitations found therein.
The prominent examples in the SC phase are the well-known phase (Leggett) and amplitude (Higgs) modes \cite{Varma02,Cea14,Measson14,Tsuji15,Krull16,Schwarz20,Fan22}.
The former is located at zero energy as our model describes neutral superfluids.
The latter is found at the lower edge of the two-particle continuum.

@@

In our study, we employ the extended Hubbard model because it encompasses various phases and thereby provides direct access to a rich spectrum of excitation spectra therein.
Its Hamiltonian is given by
\begin{equation}
    \label{eqn:full_hamiltonian}
    \begin{aligned}
        H = &-t \sum_{\langle i, j \rangle, \sigma} \left( c_{i\sigma}^\dagger c_{j\sigma} + \text{h.c.} \right) 
        + \mu \sum_{i,\sigma} n_{i\sigma} \\
        &+ U \sum_{i} n_{i\uparrow} n_{i\downarrow} 
        + \frac{V}{2} \sum_{\langle i, j\rangle, \sigma} n_{i\sigma} n_{j\sigma}\,,
    \end{aligned}
\end{equation}
where $c_{i\sigma}^{(\dagger)}$ annihilates (creates) an electron with spin $\sigma$ on lattice site $i$ 
and $\langle i, j\rangle$ denotes the summation over the next neighbors.
The parameters are the hopping amplitude $t$, the onsite interaction $U$, the intersite interaction $V$, and the chemical potential $\mu$.
Applying a Fourier transform into $k$ space yields the single-particle dispersion 
\begin{equation}
    \epsilon_0 (\vk) = -2t \sum_{\alpha=1}^D \cos(k_\alpha)\,,\quad k_\alpha \in [-\pi, \pi)\,,
\end{equation}
with the system's dimension $D$ and the dimensionless momentum $\vk$.

@@

This procedure yields the ground state phase diagram at temperature $T=0$ shown in \autoref{fig:phase_diagram}.
The CDW-AFM boundary is located at $U = zV$, where $z$ is the coordination.
This holds for both the square lattice ($z=4$) and the simple cubic lattice ($z=6$) and can be seen by comparing \eqref{eqn:delta_cdw} and \eqref{eqn:delta_afm}:
Crossing the aforementioned line, changes which one of the two prefactors is larger.

@@

In this article, we will be using the operator pseudo scalar product
\begin{equation}
\label{eqn:scalar_product}
    (A | B) \coloneqq  \langle [A^\dagger, B] \rangle\,,
\end{equation}
where the expectation values are taken with respect to the thermal equilibrium of the mean-field Hamiltonian \eqref{eqn:mf_hamiltonian}.
Note, that this is not a proper scalar product, as a norm defined using it can also be negative.
For example, assume that for some operator $A$
\begin{equation}
    ||A|| = (A | A) = \langle [A^\dagger, A] \rangle > 0\,,
\end{equation}
then
\begin{equation}
    ||A^\dagger|| = (A^\dagger | A^\dagger) = \langle [A, A^\dagger] \rangle = - ||A|| < 0\,.
\end{equation}
We will compute $\mM$ by commutating with the full Hubbard Hamiltonian \eqref{eqn:full_hamiltonian} enabling us to capture collective behavior rather than the one particle dynamics given by the mean-field approximation.

@@

Next, we order these operators in a way reminiscent of the $x$- and $p$-terms in the harmonic oscillator, i.e.,
\begin{equation}
    X_i \coloneqq  A_i + A_i^\dagger\,,\quad P_i \coloneqq  A_i - A_i^\dagger\,.
\end{equation}
Then, our set is given by $\mathcal{B}_{XP} \coloneqq \{ X_i, P_i \}$.
If some operators would be 0 ($P_i$ for any $n$-type term) or duplicates (certain $X_i$ for $g$-type terms), we omit them.

@@

To study the collective excitations, we investigate various retarded Green's functions 
\begin{equation}
    G_{AB}^\text{ret} (t) = - i \langle [A(t), B] \rangle \Theta(t)\,,
\end{equation}
and their Fourier-transformed
\begin{equation}
    \label{eqn:standard_gf}
    G_{AB}(\omega + i0^+) = -i \int_0^{\infty} e^{izt} \langle [A(t), B] \rangle \mathrm{d}t\,,
\end{equation}
where $\Theta(t)$ is the Heaviside function.
Our previous choice of pseudo scalar product enables us to write down a matrix-valued Green's functions in terms of the aforementioned matrices
\begin{align}
    \label{eqn:green_function}
    \mathcal{G}(z &= \omega + i0^+) = \mN \frac{1}{-z \mN - \mM} \mN \\
        &\coloneqq  -\mN \mathcal{R}(z) \mN\,,
\end{align}
where we used the resolvent
\begin{equation}
    \label{eqn:resolvent}
    \mathcal{R}(z) \coloneqq  \frac{1}{z \mN + \mM}\,.
\end{equation}
The entries of this matrix $\mathcal{G}(z)$ are the Fourier-transformed Green's functions mentioned in \eqref{eqn:standard_gf} with respect to the operators from the initially chosen set.
For instance, assume the operator in the set is $f_{\vk}$, then the top left element of $\mathcal{G}$ is the Green's function 
\begin{align}
    \mathcal{G}_{00}(z = \omega +i0^+) &=  G_{f_{\vk}(t) f_{\vk}^\dagger} (z) \nonumber \\
        &= -i \int_0^{\infty} \langle [f_{\vk}(t), f_{\vk}^\dagger(0)] \rangle e^{izt} \mathrm{d}t\,.
\end{align}

@@

where $a_i$ and $b_i$ are the Lanczos coefficients \cite{PettiforRecursion,ViswanathRecursion}.
Obtaining a diagonal block of $\mathcal{G}$ is easily achieved by setting $\vec{b}(0) = \vec{a}(0)$.
Obtaining an off-diagonal block requires linear combinations in the form of 
\begin{align}
    \vec{a}^\dagger \mathcal{G} \vec{b} &= \frac{1}{4} \Big\{ 
        \underbrace{\left[ (\vec{a} + \vec{b})^\dagger \mathcal{G} (\vec{a} + \vec{b}) - (\vec{a} - \vec{b})^\dagger \mathcal{G} (\vec{a} - \vec{b}) \right]}_{ = 2 \vec{a}^\dagger \mathcal{G} \vec{b} + 2 \vec{b}^\dagger \mathcal{G} \vec{a}} \\
    &-i \underbrace{\left[ (\vec{a} + i \vec{b})^\dagger \mathcal{G} (\vec{a} + i \vec{b}) - (\vec{a} - i \vec{b})^\dagger \mathcal{G} (\vec{a} - i \vec{b}) \right]}_{= 2i \vec{a}^\dagger \mathcal{G} \vec{b} - 2i \vec{b}^\dagger \mathcal{G} \vec{a}} \Big\}\,, \nonumber
\end{align}
i.e., one computes the four diagonal blocks and then simply adds them as described above.

@@

To begin, let us describe the spectral functions $\mathcal{A}_i (\omega) = - (1/\pi) \Im [\mathcal{G}_i (\omega + i0^+)]$ with respect to the operators above in the various phases.
\autoref{fig:resolvent_overview_SC} shows them at $U = -2.5t$ in the SC ($V=-0.1t$) and CDW ($V=0.1t$ and $V=0.5t$) phase, as well as in the SC-CDW coexistence at $V=0$.

There are a variety of different features. The difference between the two lattices is mostly quantitative.
First and foremost, in the SC phase, see panel (a), there is a sharp peak located at $\omega=0$ in $\spectral{Phase}$ and a singularity in $\spectral{Higgs}$ located at $\omega=2\Delta_\text{SC}$.
We identify them with the well-known Leggett and Higgs modes in superconductors.
The former can be found at zero energy because our model describes a neutral superfluid wherein there is no coupling of the phase of the order parameter to the electromagnetic potential.
Therefore, this phase may be chosen arbitrarily, and changing it requires no energy which ultimately allows a Goldstone mode to exist in the first place \cite{Goldstone1961,Anderson63}.

@@

This analysis provides that the real part of $\greens{Phase}$ near $\omega=0$ behaves as $1/\omega^2$ indicating that the peak in $\spectral{Phase}$ is the derivative of a $\delta$ distribution.
Note, that it is located at $\omega=0$ of a perfectly asymmetrical bosonic spectral function and therefore must have 0 weight.
This is given here, because $\int \delta'(\omega) \mathrm{d}\omega = 0$.

@@

At $V=0$, both the SC and the CDW orders coexist. 
In this phase, the mean-field single-particle energies are given by
\begin{equation}
    E_{\pm} = \pm \sqrt{\epsilon^2 + |\Delta_\text{CDW}|^2 + |\Delta_\text{SC}|^2}\,.
\end{equation}
This means that the proper order parameter is $\Delta_\text{tot} \coloneqq \sqrt{|\Delta_\text{CDW}|^2 + |\Delta_\text{SC}|^2}$.
Here, $\Delta_\text{CDW}$ and $\Delta_\text{SC}$ can change arbitrarily without affecting the system's energy as long as $\Delta_\text{tot}$ remains constant.
Note that the degeneracy of both phases is exact due to the SO(4) symmetry \cite{yang90}.

@@

This concludes our proof. Note, that the only assumption made, is that the expectation is to be taken with respect to the thermal equilibrium.

@@

We would like to use operators like \eqref{eqn:ieom_basis_operator} to span our operator set.
However, numerically, it is impossible to matrices with infinite dimensions, requiring us to discretize.
Additionally, we cannot numerically write down $\delta$ distributions as matrix elements. Therefore, we need some special treatments, which we will describe in this section.
We discretize the $\gamma$-space equally with each abscissa $\gamma_i$ being spaced by $\Delta \gamma$.
Then, we define an approximate $\delta$ function
\begin{equation}
    h(\gamma) \coloneqq \begin{cases}
        \frac{1}{\Delta \gamma} & |\gamma| < \frac{\Delta \gamma}{2} \\ 0 & \text{otherwise}
    \end{cases}\,.
\end{equation}
Computing an integral yields
\begin{equation}
    \int f(\tilde{\gamma}) h(\tilde{\gamma} - \gamma_j) \mathrm{d} \tilde{\gamma} \approx \sum_i f(\gamma_i) h(\gamma_i - \gamma_j) \Delta \gamma = f(\gamma_j)\,,
\end{equation}
showing that $h(\gamma)$ mimicks the continuous $\delta$ distribution for a discrete description.
Note, that if one takes the limit $\Delta \gamma \to 0$ the $\delta$ distribution is exactly reproduced.

Numerically, we can compute the matrix elements like this
\begin{equation}
    \label{eqn:numerical_example}
    (A_i | A_j) = \frac{1}{N} \sum_{\vk \vl} h(\gamma_i - \hat{\gamma}(\vk)) h(\gamma_j - \hat{\gamma}(\vl)) (A_{\vk} | A_{\vl})\,.
\end{equation}
Here, all expressions are finite numbers and do not require further treatment.
Mutatis mutandis, one obtains a similar expression for $(A_i | [H, A_j])$.
Additionally, note, that the expressions $(A_{\vk} | A_{\vl})$ and $(A_{\vk} | [H, A_{\vl}])$ can be reduced to sums of terms that are proportional to $\gamma_i$ or $\gamma_j$, respectively.
Therefore, it is not necessary to implement any sort of momentum space computations.
For example, consider $A_{\vk} = f_{\vk}^\dagger$, then
\begin{equation}
    ( f_{\vk}^\dagger | f_{\vl}^\dagger ) = \delta_{\vk, \vl}\, \langle 1 - n_{\vk \uparrow} - n_{\vk \downarrow} \rangle \eqqcolon \delta_{\vk, \vl}\, \alpha(\vk)
\end{equation}
holds. Inserting this into \eqref{eqn:numerical_example} yields
\begin{equation}
    (A_i | A_j) = \frac{1}{N} \sum_{\vk} h(\gamma_i - \hat{\gamma}(\vk)) h(\gamma_j - \hat{\gamma}(\vk)) \alpha(\vk)\,.
\end{equation}
As stated in the main text, see \eqref{eqn:equal_expecs}, $\alpha(\vk)$ depends only on $\hat{\gamma}(\vk)$, hence we can switch from the momentum sum to a $\gamma$ integration
\begin{align}
    (A_i | A_j) &= \int \rho(\gamma) h(\gamma_i - \gamma) h(\gamma_j - \gamma) \alpha(\gamma) \dgamma \nonumber \\
        &\approx \rho(\gamma_i) h(\gamma_j - \gamma_i) \alpha(\gamma_i) \nonumber \\
        &= \frac{\delta_{ij}}{\Delta \gamma} \rho(\gamma_i) \alpha(\gamma_i)\,.
\end{align}
Since $\Delta \gamma$ is a positive constant factor, it does not impact any of our matrix operations, i.e., we can pull it out to the front of the expressions $\mM \to \tilde{m} (\Delta \gamma)^2$ and $\mN \to \tilde{n} (\Delta \gamma)^2$.
Then, we can rewrite \eqref{eqn:green_derivation} as
\begin{equation}
    G_{AB^\dagger}(z) = - (\Delta \gamma)^2 \vec{b}^\dagger (0) \left[ \tilde{n} \frac{1}{\tilde{m} + z \tilde{n}} \tilde{n} \right] \vec{a}(0)\,.
\end{equation}
To give a practical example, let us compute $\greens{Higgs}$ from \eqref{eqn:resolvent_bases}.
We order our operator set such that all $f_\gamma + f_\gamma^{\dagger}$ are upfront. 
This requires the vector elements $\vec{a}$ and $\vec{b}$ to be
\begin{equation}
    a_i = b_i = \begin{cases}
        1 & i < N_\gamma \\
        0 & \text{otherwise}
    \end{cases}\,,
\end{equation}
where $N_\gamma$ is the number of discretization points.
Using this, we can expand the relation for the Green's function
\begin{align}
    \greens{Higgs} (z) &= -(\Delta \gamma)^2 \sum_{i=0}^{N_\gamma} \sum_{j=0}^{N_\gamma} \left[ \tilde{n} \frac{1}{\tilde{m} + z \tilde{n}} \tilde{n} \right]_{ij} \nonumber \\
        &\approx - \int \dgamma \int \dgamma' \left[ \tilde{n} \frac{1}{\tilde{m} + z \tilde{n}} \tilde{n} \right](\gamma, \gamma')\,.
\end{align}
The first line is an approximation of the integral in the second line, as each matrix element corresponds to a specific value for $\gamma$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                  Abstract                                                 %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
    We investigate the superconducting (SC), charge-density wave (CDW), and antiferromagnetic (AFM) phases 
    in the extended Hubbard model at zero temperature and half-filling.
    We employ the iterated equations of motion approach \cite{Kalthoff17,bleicker18} to compute the two-particle Green's functions and by extension, the corresponding spectral densities.
    This renders comprehensive analysis of the behavior of collective excitations possible as the model's parameters are changed across phase transitions.
    We identify the well-known amplitude (Higgs) and phase (Anderson-Bogoliubov) modes within the superconducting phase and observe a similar excitation in the CDW phase which shifts towards zero energy as the system approaches the phase transition to the SC phase.
    In the CDW phase, close to the phase transition to the AFM phase, we find a collective mode that does not change significantly and another mode that becomes soft as the phase boundary is approached.
\end{abstract}

@@

In this article, we will first employ a mean-field approximation to the interaction terms 
and then use the methodology from the so-called iterated equations of motion approach (iEoM),
which has already seen success in the handling of interaction quenches \cite{uhrig09,hamerla13,hamerla14,bleicker18}.
Here, one commutes operators from a suitable operator basis with the Hamiltonian and adds the newly occurring terms to it.
Naturally, a truncation is necessary for most practical applications, however, the approximation becomes better the more terms are included within the basis.
The functionality of this method has been compared to the density matrix formalism and has proved to be considerably more accurate \cite{Kalthoff17}.
Additionally, we will show an explicit way to obtain various Green's functions from the aforementioned methodology.
By extension, we also obtain the spectral functions of the investigated systems and discuss the excitations found therein.
The prominent examples in the SC phase are the well-known phase (Anderson-Bogoliubov) and amplitude (Higgs) modes \cite{Kulik1981,Varma02,Cea14,Measson14,Tsuji15,Krull16,Schwarz20,Fan22}.
The former is located at zero energy as our model describes neutral superfluids.
The latter is found at the lower edge of the two-particle continuum.

@@

In our study, we employ the extended Hubbard model as it encompasses assorted phases and thereby provides direct access to the rich excitation spectra therein.
Its Hamiltonian is given by
\begin{equation}
    \label{eqn:full_hamiltonian}
    \begin{aligned}
        H = &-t \sum_{\langle i, j \rangle, \sigma} \left( c_{i\sigma}^\dagger c_{j\sigma} + \text{h.c.} \right) 
        + \mu \sum_{i,\sigma} n_{i\sigma} \\
        &+ U \sum_{i} n_{i\uparrow} n_{i\downarrow} 
        + \frac{V}{2} \sum_{\langle i, j\rangle, \sigma} n_{i\sigma} n_{j\sigma}\,,
    \end{aligned}
\end{equation}
where $c_{i\sigma}^{(\dagger)}$ annihilates (creates) an electron with spin $\sigma$ on lattice site $i$ 
and $\langle i, j\rangle$ denotes the summation over the next neighbors.
The parameters are the hopping amplitude $t$, the onsite interaction $U$, the intersite interaction $V$, and the chemical potential $\mu$.
Applying a Fourier transform into $k$ space yields the single-particle dispersion 
\begin{equation}
    \epsilon_0 (\vk) = -2t \sum_{\alpha=1}^D \cos(k_\alpha)\,,\quad k_\alpha \in [-\pi, \pi)\,,
\end{equation}
with the system's dimension $D$ and the dimensionless momentum $\vk$.

@@

This procedure yields the ground state phase diagram at temperature $T=0$ shown in \autoref{fig:phase_diagram}.
The CDW-AFM boundary is located at $U = zV$, where $z$ is the coordination.
This holds for both the square lattice ($z=4$) and the simple cubic lattice ($z=6$) and can be seen by comparing \eqref{eqn:delta_cdw} and \eqref{eqn:delta_afm}:
Crossing the aforementioned line, changes which prefactor of the two order parameters is larger.

@@

In this article, we will be using the operator pseudo scalar product
\begin{equation}
\label{eqn:scalar_product}
    (A | B) \coloneqq  \langle [A^\dagger, B] \rangle\,,
\end{equation}
where the expectation values are taken with respect to the thermal equilibrium of the mean-field Hamiltonian \eqref{eqn:mf_hamiltonian}.
Note, that this is not a proper scalar product, as a norm defined using it can also be negative.
For example, assume that for some operator $A$
\begin{equation}
    ||A|| = (A | A) = \langle [A^\dagger, A] \rangle > 0\,,
\end{equation}
then
\begin{equation}
    ||A^\dagger|| = (A^\dagger | A^\dagger) = \langle [A, A^\dagger] \rangle = - ||A|| < 0\,.
\end{equation}
Nevertheless, this does not pose any issue in our calculations.
We will compute $\mM$ by commutating with the full Hubbard Hamiltonian \eqref{eqn:full_hamiltonian} enabling us to capture collective behavior rather than the one particle dynamics given by the mean-field approximation.

@@

Next, we order these operators in a way reminiscent of the $x$- and $p$-terms in the harmonic oscillator, i.e.,
\begin{equation}
    X_i \coloneqq  A_i + A_i^\dagger\,,\quad P_i \coloneqq  A_i - A_i^\dagger\,.
\end{equation}
Then, our set is given by $\mathcal{B}_{XP} \coloneqq \{ X_i, \ldots, P_i, \ldots \}$.
If some operators would be 0 ($P_i$ for any $n$-type term) or duplicates (certain $X_i$ for $g$-type terms), we omit them.

@@

To study the collective excitations, we investigate various retarded Green's functions 
\begin{equation}
    G_{AB}^\text{ret} (t) = - i \langle [A(t), B] \rangle \Theta(t)\,,
\end{equation}
and their Fourier-transformed
\begin{equation}
    \label{eqn:standard_gf}
    G_{AB}(z = \omega + i0^+) = -i \int_0^{\infty} e^{izt} \langle [A(t), B] \rangle \mathrm{d}t\,,
\end{equation}
where $\Theta(t)$ is the Heaviside function.
Our previous choice of pseudo scalar product enables us to write down a matrix-valued Green's functions in terms of the aforementioned matrices
\begin{align}
    \label{eqn:green_function}
    \mathcal{G}(z &= \omega + i0^+) = \mN \frac{1}{-z \mN - \mM} \mN \\
        &\coloneqq  -\mN \mathcal{R}(z) \mN\,,
\end{align}
where we used the resolvent
\begin{equation}
    \label{eqn:resolvent}
    \mathcal{R}(z) \coloneqq  \frac{1}{z \mN + \mM}\,.
\end{equation}
The entries of this matrix $\mathcal{G}(z)$ are the Fourier-transformed Green's functions mentioned in \eqref{eqn:standard_gf} with respect to the operators from the initially chosen set.
For instance, assume the operator in the set is $f_{\vk}$, then the top left element of $\mathcal{G}$ is the Green's function 
\begin{align}
    \mathcal{G}_{00}(z = \omega +i0^+) &=  G_{f_{\vk} f_{\vk}^\dagger} (z) \nonumber \\
        &= -i \int_0^{\infty} \langle [f_{\vk}(t), f_{\vk}^\dagger(0)] \rangle e^{izt} \mathrm{d}t\,.
\end{align}

@@

where $a_i$ and $b_i$ are the Lanczos coefficients \cite{PettiforRecursion,ViswanathRecursion}.

Obtaining a diagonal block of $\mathcal{G}$ is easily achieved by setting $\vec{b}(0) = \vec{a}(0)$.
Obtaining an off-diagonal block requires linear combinations in the form of 
\begin{align}
    \vec{a}^\dagger \mathcal{G} \vec{b} &= \frac{1}{4} \Big\{ 
        \underbrace{\left[ (\vec{a} + \vec{b})^\dagger \mathcal{G} (\vec{a} + \vec{b}) - (\vec{a} - \vec{b})^\dagger \mathcal{G} (\vec{a} - \vec{b}) \right]}_{ = 2 \vec{a}^\dagger \mathcal{G} \vec{b} + 2 \vec{b}^\dagger \mathcal{G} \vec{a}} \\
    &-i \underbrace{\left[ (\vec{a} + i \vec{b})^\dagger \mathcal{G} (\vec{a} + i \vec{b}) - (\vec{a} - i \vec{b})^\dagger \mathcal{G} (\vec{a} - i \vec{b}) \right]}_{= 2i \vec{a}^\dagger \mathcal{G} \vec{b} - 2i \vec{b}^\dagger \mathcal{G} \vec{a}} \Big\}\,, \nonumber
\end{align}
i.e., one computes the four diagonal blocks and then simply adds them as described above.

@@

To begin, let us describe the spectral functions $\mathcal{A}_i (\omega) = - (1/\pi) \Im [\mathcal{G}_i (\omega + i0^+)]$ with respect to the operators above in the various phases.
\autoref{fig:resolvent_overview_SC} shows them at $U = -2.5t$ in the SC ($V=-0.1t$) and CDW ($V=0.1t$ and $V=0.5t$) phase.

There are a variety of different features. The difference between the two lattices is mostly quantitative.
First and foremost, in the SC phase, see panel (a), 
there is a sharp peak located at $\omega=0$ in $\spectral{Phase}$ and a singularity in $\spectral{Higgs}$ located at $\omega=2\Delta_\text{SC}$.
We identify them with the well-known Anderson-Bogoliubov and Higgs modes in superconductors.
Note, that the former must not have any weight though, because the spectral function is perfectly antisymmetrical.
The peak itself can be found at zero energy because our model describes a neutral superfluid 
wherein there is no coupling of the phase of the order parameter to the electromagnetic potential.
Therefore, this phase may be chosen arbitrarily, and changing it requires no energy 
which ultimately allows a Goldstone mode to exist in the first place \cite{Goldstone1961,Anderson63}.

@@

This analysis provides that the real part of $\greens{Phase}$ near $\omega=0$ behaves as $1/\omega^2$ indicating that the peak in $\spectral{Phase}$ is the derivative of a $\delta$ distribution.
The weight of such a peak is given by the integral $\int \delta'(\omega) \mathrm{d}\omega = 0$.
This confirms our previous expectation that this peak cannot have any weight.

@@

At $V=0$, both the SC and the CDW orders coexist. 
In this phase, the mean-field single-particle energies are given by
\begin{equation}
    E_{\pm} = \pm \sqrt{\epsilon^2 + |\Delta_\text{CDW}|^2 + |\Delta_\text{SC}|^2}\,.
\end{equation}
This means that the proper order parameter is $\Delta_\text{tot} \coloneqq \sqrt{|\Delta_\text{CDW}|^2 + |\Delta_\text{SC}|^2}$.
Here, $\Delta_\text{CDW}$ and $\Delta_\text{SC}$ can change arbitrarily without affecting the system's energy as long as $\Delta_\text{tot}$ remains constant.
Note, that the degeneracy of both phases is exact due to the SO(4) symmetry \cite{yang90}.

@@

This concludes our proof. Note, that the only assumption made, is that the expectation value is to be taken with respect to the thermal equilibrium.

@@

We would like to use operators like \eqref{eqn:ieom_basis_operator} to span our operator set.
However, numerically, it is impossible to matrices with infinite dimensions, requiring us to discretize.
Additionally, we cannot numerically write down $\delta$ distributions as matrix elements. Therefore, we need some special treatments, which we will describe in this section.
We discretize the $\gamma$-space equally with each abscissa $\gamma_i$ being spaced by $\Delta \gamma$.
Then, we define an approximate $\delta$ function
\begin{equation}
    h(\gamma) \coloneqq \begin{cases}
        \frac{1}{\Delta \gamma} & |\gamma| < \frac{\Delta \gamma}{2} \\ 0 & \text{otherwise}
    \end{cases}\,.
\end{equation}
Computing an integral numerically yields
\begin{equation}
    \int f(\tilde{\gamma}) h(\tilde{\gamma} - \gamma_j) \mathrm{d} \tilde{\gamma} \approx \sum_i f(\gamma_i) h(\gamma_i - \gamma_j) \Delta \gamma = f(\gamma_j)\,,
\end{equation}
showing that $h(\gamma)$ mimicks the continuous $\delta$ distribution for a discrete description.
Note, that if one takes the limit $\Delta \gamma \to 0$ the $\delta$ distribution is exactly reproduced.

Numerically, we can compute the matrix elements like this
\begin{equation}
    \label{eqn:numerical_example}
    (A_i | A_j) = \frac{1}{N} \sum_{\vk \vl} h(\gamma_i - \hat{\gamma}(\vk)) h(\gamma_j - \hat{\gamma}(\vl)) (A_{\vk} | A_{\vl})\,.
\end{equation}
Here, all expressions are finite numbers and do not require further treatment.
Mutatis mutandis, one obtains a similar expression for $(A_i | [H, A_j])$.
Additionally, note, that the expressions $(A_{\vk} | A_{\vl})$ and $(A_{\vk} | [H, A_{\vl}])$ can be reduced to sums of terms that are proportional to $\gamma_i$ or $\gamma_j$, respectively.
Therefore, it is not necessary to implement any sort of momentum space computations.
For example, consider $A_{\vk} = f_{\vk}^\dagger$, then
\begin{equation}
    ( f_{\vk}^\dagger | f_{\vl}^\dagger ) = \delta_{\vk, \vl}\, \langle 1 - n_{\vk \uparrow} - n_{\vk \downarrow} \rangle \eqqcolon \delta_{\vk, \vl}\, \alpha(\vk)
\end{equation}
holds. Inserting this into \eqref{eqn:numerical_example} yields
\begin{equation}
    (A_i | A_j) = \frac{1}{N} \sum_{\vk} h(\gamma_i - \hat{\gamma}(\vk)) h(\gamma_j - \hat{\gamma}(\vk)) \alpha(\vk)\,.
\end{equation}
As stated in the main text, see \eqref{eqn:equal_expecs}, $\alpha(\vk)$ depends only on $\hat{\gamma}(\vk)$, hence we can switch from the momentum sum to a $\gamma$ integration
\begin{align}
    (A_i | A_j) &= \int \rho(\gamma) h(\gamma_i - \gamma) h(\gamma_j - \gamma) \alpha(\gamma) \dgamma \nonumber \\
        &\approx \rho(\gamma_i) h(\gamma_j - \gamma_i) \alpha(\gamma_i) \nonumber \\
        &= \frac{\delta_{ij}}{\Delta \gamma} \rho(\gamma_i) \alpha(\gamma_i)\,.
\end{align}
Since $\Delta \gamma$ is a positive constant factor, it does not impact any of our matrix operations, i.e., we can pull it out to the front of the expressions $\mM \to \tilde{m} (\Delta \gamma)^2$ and $\mN \to \tilde{n} (\Delta \gamma)^2$.
Then, we can rewrite \eqref{eqn:green_derivation} as
\begin{equation}
    G_{AB^\dagger}(z) = - (\Delta \gamma)^2 \vec{b}^\dagger (0) \left[ \tilde{n} \frac{1}{\tilde{m} + z \tilde{n}} \tilde{n} \right] \vec{a}(0)\,.
\end{equation}

To give a practical example, let us compute $\greens{Higgs}$ from \eqref{eqn:resolvent_bases}.
We order our operator set such that all $f_\gamma + f_\gamma^{\dagger}$ are upfront. 
This requires the vector elements $\vec{a}$ and $\vec{b}$ to be
\begin{equation}
    a_i = b_i = \begin{cases}
        1 & i < N_\gamma \\
        0 & \text{otherwise}
    \end{cases}\,,
\end{equation}
where $N_\gamma$ is the number of discretization points.
Using this, we can expand the relation for the Green's function
\begin{align}
    \greens{Higgs} (z) &= -(\Delta \gamma)^2 \sum_{i=0}^{N_\gamma} \sum_{j=0}^{N_\gamma} \left[ \tilde{n} \frac{1}{\tilde{m} + z \tilde{n}} \tilde{n} \right]_{ij} \nonumber \\
        &\approx - \int \dgamma \int \dgamma' \left[ \tilde{n} \frac{1}{\tilde{m} + z \tilde{n}} \tilde{n} \right](\gamma, \gamma')\,.
\end{align}
The first line is an approximation of the integral in the second line, as each matrix element corresponds to a specific value for $\gamma$.
